PANCHANGE_data_preprocessing

### 2021-01-11 PANCHANGE_analysis.Rmd
detail checks and fixes

```{r model1b education function}
model1b.estimates <- function(outcome, data_set) {
  # define the variable name of the dependent variable
  dependent.variable <- outcome
  # define the independent variables
  independent.variables <- c("depression_and_anxiety",
                             "eating_disorders",
                             "obsessive_compulsive_disorders",
                             "bipolar_and_schizophrenia",
                             "mhd.posttraumatic_stress_disorder_ptsd",
                             "autism_spectrum_disorder",
                             "mhd.attention_deficit_hyperactivity_disorder",
                             "mhd.personality_disorder",
                             "Gender_collapsed",
                             "age_category_collapsed",
                             "Ethnicity_collapsed",
                             "key_worker",
                             "time_diff_coping_lockdown_march_23",
                             "highest_education"
                                )
  # create the formula
  glm.formula <- as.formula(paste(dependent.variable, paste(independent.variables, collapse=" + "), sep=" ~ "))
  glm.formula

@@

@@

```{r model1b.explan education function}
model1b.explan.estimates <- function(outcome, explan_var, data_set) {
  # define the variable name of the dependent variable
  dependent.variable <- outcome
  # define the independent variables
  independent.variables <- c(explan_var,
                             "depression_and_anxiety",
                             "eating_disorders",
                             "obsessive_compulsive_disorders",
                             "bipolar_and_schizophrenia",
                             "mhd.posttraumatic_stress_disorder_ptsd",
                             "autism_spectrum_disorder",
                             "mhd.attention_deficit_hyperactivity_disorder",
                             "mhd.personality_disorder",
                             "Gender_collapsed",
                             "age_category_collapsed",
                             "Ethnicity_collapsed",
                             "key_worker",
                             "time_diff_coping_lockdown_march_23",
                             "highest_education"
  )
  # create the formula
  glm.formula <- as.formula(paste(dependent.variable, paste(independent.variables, collapse=" + "), sep=" ~ "))
  glm.formula

@@

```{r model1b.explan employment prior covid function}
model1b.explan.estimates <- function(outcome, explan_var, data_set) {
  # define the variable name of the dependent variable
  dependent.variable <- outcome
  # define the independent variables
  independent.variables <- c(explan_var,
                             "depression_and_anxiety",
                             "eating_disorders",
                             "obsessive_compulsive_disorders",
                             "bipolar_and_schizophrenia",
                             "mhd.posttraumatic_stress_disorder_ptsd",
                             "autism_spectrum_disorder",
                             "mhd.attention_deficit_hyperactivity_disorder",
                             "mhd.personality_disorder",
                             "Gender_collapsed",
                             "age_category_collapsed",
                             "Ethnicity_collapsed",
                             "time_diff_coping_lockdown_march_23",
                             "employment_prior_covid"
  )
  # create the formula
  glm.formula <- as.formula(paste(dependent.variable, paste(independent.variables, collapse=" + "), sep=" ~ "))
  glm.formula

@@

```{r model1b.explan employment change function}
model1b.explan.estimates <- function(outcome, explan_var, data_set) {
  # define the variable name of the dependent variable
  dependent.variable <- outcome
  # define the independent variables
  independent.variables <- c(explan_var,
                             "depression_and_anxiety",
                             "eating_disorders",
                             "obsessive_compulsive_disorders",
                             "bipolar_and_schizophrenia",
                             "mhd.posttraumatic_stress_disorder_ptsd",
                             "autism_spectrum_disorder",
                             "mhd.attention_deficit_hyperactivity_disorder",
                             "mhd.personality_disorder",
                             "Gender_collapsed",
                             "age_category_collapsed",
                             "Ethnicity_collapsed",
                             "key_worker",
                             "time_diff_coping_lockdown_march_23",
                             "employment_change"
  )
  # create the formula
  glm.formula <- as.formula(paste(dependent.variable, paste(independent.variables, collapse=" + "), sep=" ~ "))
  glm.formula

@@

@@

@@

# GAD prepan
### Model 1b: Overlapping disorders, Adjusted
Model 0a + Gender_collapsed + age_category_collapsed + Ethnicity_collapsed + Time registration and baseline
```{r gad employment change sensitivity model1b sum_score prepan}
# Regression coefficients
gad_prepan_model1b_education_estimates <- model1b.explan.estimates(outcome = "gad.sum_score_base",
                         explan_var = "pcl.sum_score_prepan",
                         data_set = "dat.prepan")
kable(gad_prepan_model1b_education_estimates)
gad_prepan_model1b_education_estimates

# PHQ baseline
### Model 1b: Overlapping disorders, Adjusted
Model 0a + Gender_collapsed + age_category_collapsed + Ethnicity_collapsed + Time registration and baseline
```{r phq employment change sensitivity model1b absolute whole sample}
# Regression coefficients
phq_baseline_model1b_estimates <- model1b.estimates(outcome = "phq.sum_score_base",
                                                 data_set = "dat")
kable(phq_baseline_model1b_estimates)
phq_baseline_model1b_estimates

@@

# PHQ prepan
### Model 1b: Overlapping disorders, Adjusted
Model 0a + Gender_collapsed + age_category_collapsed + Ethnicity_collapsed + Time registration and baseline
```{r phq employment change sensitivity model1b sum_score prepan}
# Regression coefficients
phq_prepan_model1b_education_estimates <- model1b.explan.estimates(outcome = "phq.sum_score_base",
                         explan_var = "pcl.sum_score_prepan",
                         data_set = "dat.prepan")
kable(phq_prepan_model1b_education_estimates)
phq_prepan_model1b_education_estimates

@@

@@

@@

@@

@@

@@

@@

@@

```{r run models for rero by sample forest plots}

@@

@@

@@

@@

>>>>

```{r model1b education function}
model1b.estimates <- function(outcome, data_set) {
  # define the variable name of the dependent variable
  dependent.variable <- outcome
  # define the independent variables
  independent.variables <- c("depression_and_anxiety",
                             "eating_disorders",
                             "obsessive_compulsive_disorders",
                             "bipolar_and_schizophrenia",
                             "mhd.posttraumatic_stress_disorder_ptsd",
                             "autism_spectrum_disorder",
                             "mhd.attention_deficit_hyperactivity_disorder",
                             "mhd.personality_disorder",
                             "Gender_collapsed",
                             "age_category_collapsed_reordered",
                             "Ethnicity_collapsed",
                             "key_worker",
                             "time_diff_coping_lockdown_march_23",
                             "highest_education"
                                )
  # create the formula
  glm.formula <- as.formula(paste(dependent.variable, paste(independent.variables, collapse=" + "), sep=" ~ "))
  glm.formula

@@

                "age_category_collapsed16-18" = "16-18 years",

@@

```{r model1b.explan education function}
model1b.explan.estimates <- function(outcome, explan_var, data_set) {
  # define the variable name of the dependent variable
  dependent.variable <- outcome
  # define the independent variables
  independent.variables <- c(explan_var,
                             "depression_and_anxiety",
                             "eating_disorders",
                             "obsessive_compulsive_disorders",
                             "bipolar_and_schizophrenia",
                             "mhd.posttraumatic_stress_disorder_ptsd",
                             "autism_spectrum_disorder",
                             "mhd.attention_deficit_hyperactivity_disorder",
                             "mhd.personality_disorder",
                             "Gender_collapsed",
                             "age_category_collapsed_reordered",
                             "Ethnicity_collapsed",
                             "key_worker",
                             "time_diff_coping_lockdown_march_23",
                             "highest_education"
  )
  # create the formula
  glm.formula <- as.formula(paste(dependent.variable, paste(independent.variables, collapse=" + "), sep=" ~ "))
  glm.formula

@@

```{r model1b.explan employment prior covid function}
model1b.explan.estimates <- function(outcome, explan_var, data_set) {
  # define the variable name of the dependent variable
  dependent.variable <- outcome
  # define the independent variables
  independent.variables <- c(explan_var,
                             "depression_and_anxiety",
                             "eating_disorders",
                             "obsessive_compulsive_disorders",
                             "bipolar_and_schizophrenia",
                             "mhd.posttraumatic_stress_disorder_ptsd",
                             "autism_spectrum_disorder",
                             "mhd.attention_deficit_hyperactivity_disorder",
                             "mhd.personality_disorder",
                             "Gender_collapsed",
                             "age_category_collapsed_reordered",
                             "Ethnicity_collapsed",
                             "time_diff_coping_lockdown_march_23",
                             "employment_prior_covid"
  )
  # create the formula
  glm.formula <- as.formula(paste(dependent.variable, paste(independent.variables, collapse=" + "), sep=" ~ "))
  glm.formula

@@

```{r model1b.explan employment change function}
model1b.explan.estimates <- function(outcome, explan_var, data_set) {
  # define the variable name of the dependent variable
  dependent.variable <- outcome
  # define the independent variables
  independent.variables <- c(explan_var,
                             "depression_and_anxiety",
                             "eating_disorders",
                             "obsessive_compulsive_disorders",
                             "bipolar_and_schizophrenia",
                             "mhd.posttraumatic_stress_disorder_ptsd",
                             "autism_spectrum_disorder",
                             "mhd.attention_deficit_hyperactivity_disorder",
                             "mhd.personality_disorder",
                             "Gender_collapsed",
                             "age_category_collapsed_reordered",
                             "Ethnicity_collapsed",
                             "employment_prior_covid",
                             "time_diff_coping_lockdown_march_23"
  )
  # create the formula
  glm.formula <- as.formula(paste(dependent.variable, paste(independent.variables, collapse=" + "), sep=" ~ "))
  glm.formula

@@

  assign(paste0(outcome,".",explan_var,"model1b.explan_reg_stats"),
         glance(get(paste0(outcome,".",explan_var,".model1b.explan"))) %>%
           mutate(
             "n" = format(
               round(nobs,
                     digits = 0),
               nsmall = 0),
             "R2" = format(
               round(r.squared,
                     digits = 2),
               nsmall = 2),
             "adjust. R2" = format(
               round(adj.r.squared,
                     digits = 2),
               nsmall = 2),
             "F" = format(
               round(statistic,
                     digits = 2),
               nsmall = 2),
             "p value" = case_when(
               p.value < 0.001 ~ formatC(p.value, format = "e", digits = 2),
               p.value >= 0.001 & p.value < 0.01 ~ formatC(p.value, format = "f", digits = 3),
               p.value >= 0.01 ~ formatC(p.value, format = "f", digits = 2)
             )
           ) %>%
           select(model1b.estimates,
             "n",
             "R2",
             "adjust. R2",
             "F",
             "p value",
             "df"
           )
  )

@@

# Regression diagnostic statistics
# gad_baseline_model1b_reg_stats <- model1b.reg.stats(outcome = "gad.sum_score_base",
#                                                  data_set = "dat")
# kable(gad_baseline_model1b_reg_stats)
# gad_baseline_model1b_reg_stats
```

@@

# Regression diagnostic statistics
# gad_retro_model1b_explan_reg_stats <- model1b.explan.reg.stats(outcome = "gad.sum_score_base",
#                                                                explan_var = "gad.sum_score_retro",
#                                                                data_set = "dat.retro")
# kable(gad_retro_model1b_explan_reg_stats)
# gad_retro_model1b_explan_reg_stats
```

# GAD prepan
### Model 1b: Overlapping disorders, Adjusted
Model 0a + Gender_collapsed + age_category_collapsed + Ethnicity_collapsed + Time registration and baseline
```{r gad employment change sensitivity model1b sum_score prepan}
# Regression coefficients
gad_prepan_model1b_education_estimates <- model1b.explan.estimates(outcome = "gad.sum_score_base",
                         explan_var = "gad.sum_score_prepan",
                         data_set = "dat.prepan")
kable(gad_prepan_model1b_education_estimates)
gad_prepan_model1b_education_estimates
# 
# # Regression diagnostic statistics
# gad_prepan_model1b_education_reg_stats <- model1b.explan.reg.stats(outcome = "gad.sum_score_base",
#                          explan_var = "gad.sum_score_prepan",
#                          data_set = "dat.prepan")
# kable(gad_prepan_model1b_education_reg_stats)
# gad_prepan_model1b_education_reg_stats
```

# PHQ baseline
### Model 1b: Overlapping disorders, Adjusted
Model 0a + Gender_collapsed + age_category_collapsed + Ethnicity_collapsed + Time registration and baseline
```{r phq employment change sensitivity model1b absolute whole sample}
# Regression coefficients
phq_baseline_model1b_estimates <- model1b.estimates(outcome = "phq.sum_score_base",
                                                 data_set = "dat")
kable(phq_baseline_model1b_estimates)
phq_baseline_model1b_estimates
# 
# # Regression diagnostic statistics
# phq_baseline_model1b_reg_stats <- model1b.reg.stats(outcome = "phq.sum_score_base",
#                                                  data_set = "dat")
# kable(phq_baseline_model1b_reg_stats)
# phq_baseline_model1b_reg_stats
```

@@

# Regression diagnostic statistics
# phq_retro_model1b_explan_reg_stats <- model1b.explan.reg.stats(outcome = "phq.sum_score_base",
#                                                                explan_var = "phq.sum_score_retro",
#                                                                data_set = "dat.retro")
# kable(phq_retro_model1b_explan_reg_stats)
# phq_retro_model1b_explan_reg_stats
```

# PHQ prepan
### Model 1b: Overlapping disorders, Adjusted
Model 0a + Gender_collapsed + age_category_collapsed + Ethnicity_collapsed + Time registration and baseline
```{r phq employment change sensitivity model1b sum_score prepan}
# Regression coefficients
phq_prepan_model1b_education_estimates <- model1b.explan.estimates(outcome = "phq.sum_score_base",
                         explan_var = "phq.sum_score_prepan",
                         data_set = "dat.prepan")
kable(phq_prepan_model1b_education_estimates)
phq_prepan_model1b_education_estimates

# Regression diagnostic statistics
# phq_prepan_model1b_education_reg_stats <- model1b.explan.reg.stats(outcome = "phq.sum_score_base",
#                          explan_var = "phq.sum_score_prepan",
#                          data_set = "dat.prepan")
# kable(phq_prepan_model1b_education_reg_stats)
# phq_prepan_model1b_education_reg_stats
```

@@

# # Regression diagnostic statistics
# ocir_baseline_model1b_reg_stats <- model1b.reg.stats(outcome = "ocir.sum_score_base",
#                                                  data_set = "dat")
# kable(ocir_baseline_model1b_reg_stats)
# ocir_baseline_model1b_reg_stats
```

@@

# # Regression diagnostic statistics
# ocir_retro_model1b_explan_reg_stats <- model1b.explan.reg.stats(outcome = "ocir.sum_score_base",
#                                                                explan_var = "ocir.sum_score_retro",
#                                                                data_set = "dat.retro")
# kable(ocir_retro_model1b_explan_reg_stats)
# ocir_retro_model1b_explan_reg_stats
```

@@

# # Regression diagnostic statistics
# ocir_prepan_model1b_education_reg_stats <- model1b.explan.reg.stats(outcome = "ocir.sum_score_base",
#                          explan_var = "pcl.sum_score_prepan",
#                          data_set = "dat.prepan")
# kable(ocir_prepan_model1b_education_reg_stats)
# ocir_prepan_model1b_education_reg_stats
```

@@

# # Regression diagnostic statistics
# pcl_baseline_model1b_reg_stats <- model1b.reg.stats(outcome = "pcl.sum_score_base",
#                                                  data_set = "dat")
# kable(pcl_baseline_model1b_reg_stats)
# pcl_baseline_model1b_reg_stats
```

@@

# # Regression diagnostic statistics
# pcl_prepan_model1b_education_reg_stats <- model1b.explan.reg.stats(outcome = "pcl.sum_score_base",
#                          explan_var = "pcl.sum_score_prepan",
#                          data_set = "dat.prepan")
# kable(pcl_prepan_model1b_education_reg_stats)
# pcl_prepan_model1b_education_reg_stats
```

@@

```{r functions required in book for globa assignments no explan var}
lmoutput_format_noexplan <- function(lmobject,  group, dfname){

  ## Row labels
  independent.variables <- c("intercept",
                             "Comorbid depression and anxiety",
                             "Depressive disorders",
                             "Anxiety disorders",
                             "Eating disorders",
                             "Obsessive compulsive disorders",
                             "Comorbid psychotic and bipolar disorders",
                             "Psychotic disorders",
                             "Bipolar disorder",
                             "Post traumatic stress disorder",
                             "Autism spectrum disorder",
                             "Attention deficit hyperactivity disorder",
                             "Personality disorders",
                             "Female",
                             "Non-binary/Self-defined",
                             "16-18",
                             "19-25",
                             "36-45",
                             "46-55",
                             "56-65",
                             "66-70",
                             "71-75",
                             "76+",
                             "Mixed or multiple ethnicities",
                             "Asian",
                             "Black",
                             "Other",
                             "Key worker",
                             "Unemployed",
                             "Student",
                             "Retired",
                             "Time between lockdown and baseline")
  ## categories to facet on  
  categories <- c("Intercept",
                  "Pre-existing mental health (Reference controls)",
                  "Pre-existing mental health (Reference controls)",
                  "Pre-existing mental health (Reference controls)",
                  "Pre-existing mental health (Reference controls)",
                  "Pre-existing mental health (Reference controls)",
                  "Pre-existing mental health (Reference controls)",
                  "Pre-existing mental health (Reference controls)",
                  "Pre-existing mental health (Reference controls)",
                  "Pre-existing mental health (Reference controls)",
                  "Pre-existing mental health (Reference controls)",
                  "Pre-existing mental health (Reference controls)",
                  "Pre-existing mental health (Reference controls)",
                  "Gender (Reference male)",
                  "Gender (Reference male)",
                  "Age (Reference 26-35)",
                  "Age (Reference 26-35)",
                  "Age (Reference 26-35)",
                  "Age (Reference 26-35)",
                  "Age (Reference 26-35)",
                  "Age (Reference 26-35)",
                  "Age (Reference 26-35)",
                  "Age (Reference 26-35)",
                  "Ethnicity (Reference White)",
                  "Ethnicity (Reference White)",
                  "Ethnicity (Reference White)",
                  "Ethnicity (Reference White)",
                  "Employment (Reference employed)",
                  "Employment (Reference employed)",
                  "Employment (Reference employed)",
                  "Employment (Reference employed)",
                  "Other")

  ## Disorder dimension  
  groups <- c(group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group,
              group)

model1b.lm.form.noexplan <- function(outcome) {

  # define the variable name of the dependent variable
  dependent.variable <- outcome
  # define the independent variables
  independent.variables <- c("depression_and_anxiety",
                             "eating_disorders",
                             "obsessive_compulsive_disorders",
                             "bipolar_and_schizophrenia",
                             "mhd.posttraumatic_stress_disorder_ptsd",
                             "autism_spectrum_disorder",
                             "mhd.attention_deficit_hyperactivity_disorder",
                             "mhd.personality_disorder",
                             "Gender_collapsed",
                             "age_category_collapsed_reordered",
                             "Ethnicity_collapsed",
                             "employment_prior_covid",
                             "time_diff_coping_lockdown_march_23"
  )
  # create the formula
  glm.formula <- paste(dependent.variable, paste(independent.variables, collapse=" + "), sep=" ~ ")

  # assign it globally with the right variable name.
  assign(paste0(outcome,".",  "glmformula"),
         glm.formula,
         envir = .GlobalEnv

@@

### no explan var

```{r create formulas for forest models no explan}

model1b.lm.form.noexplan("gad.sum_score_base")

model1b.lm.form.noexplan("phq.sum_score_base")

model1b.lm.form.noexplan("pcl.sum_score_base")

```{r run models for all forest plot items no explan}

gad.lm.noexplan <-  lm(formula = gad.sum_score_base.glmformula,
                            dat.prepan)

phq.lm.noexplan <- lm(formula = phq.sum_score_base.glmformula,
                            dat.retro)

pcl.lm.noexplan <-  lm(formula = pcl.sum_score_base.glmformula,
                            dat.prepan)

@@

```{r run models for retro by sample forest plots}

@@

### no explanatory variables
```{r generate datasets for all plots no explan}

# grouped by disorder measure
lmoutput_format_noexplan(gad.lm.noexplan,"GAD-7", "gad.lm.for.plot")
lmoutput_format_noexplan(phq.lm.noexplan,"PHQ-9", "phq.lm.for.plot")
lmoutput_format_noexplan(pcl.lm.noexplan,"PCL", "pcl.lm.for.plot")

# Grouped by sample
lmoutput_format_noexplan(gad.lm.noexplan,"GLAD and EDGI", "gad.lm.for.plot.GLAD")
lmoutput_format_noexplan(gad.lm.noexplan,"NBR", "gad.lm.for.plot.NBR")
lmoutput_format_noexplan(gad.lm.noexplan,"RAMP", "gad.lm.for.plot.RAMP")

lmoutput_format_noexplan(phq.lm.noexplan,"GLAD and EDGI", "phq.lm.for.plot.GLAD")
lmoutput_format_noexplan(phq.lm.noexplan,"NBR", "phq.lm.for.plot.NBR")
lmoutput_format_noexplan(phq.lm.noexplan,"RAMP", "phq.lm.for.plot.RAMP")

lmoutput_format_noexplan(pcl.lm.noexplan,"GLAD and EDGI", "pcl.lm.for.plot.GLAD")
lmoutput_format_noexplan(pcl.lm.noexplan,"NBR", "pcl.lm.for.plot.NBR")
lmoutput_format_noexplan(pcl.lm.noexplan,"RAMP", "pcl.lm.for.plot.RAMP")

@@

With no control for retro or prepan (no explanatory variables)

```{r merge datasets for different combinations no explan}
# Disorder groups
plot.dat.disorder <- rbind(gad.lm.for.plot,
                                  phq.lm.for.plot,
                                  pcl.lm.for.plot)

# by sample

plot.dat.noexplan.sample.gad <- rbind(gad.lm.for.plot.GLAD,
                             gad.lm.for.plot.NBR,
                             gad.lm.for.plot.RAMP)

plot.dat.noexplan.sample.phq <- rbind(phq.lm.for.plot.GLAD,
                             phq.lm.for.plot.NBR,
                             phq.lm.for.plot.RAMP)

plot.dat.noexplan.sample.pcl <- rbind(pcl.lm.for.plot.GLAD,
                             pcl.lm.for.plot.NBR,
                             pcl.lm.for.plot.RAMP)

@@

```
### disorder (no explan)

```{r forest plot by disorder no explan, fig.height=30, fig.width=50, dpi=120}

lm.forest(plot.dat.disorder,"Predictors of pandemic symptoms")

@@

### Sample pandemic only, no explanatory variable)

```{r forest plot by sample no explan gad, fig.height=30, fig.width=50, dpi=120}

lm.forest(plot.dat.noexplan.sample.gad  ,"Predictors of anxiety across samples")

```{r forest plot by sample no explan phq, fig.height=30, fig.width=50, dpi=120}

lm.forest(plot.dat.noexplan.sample.phq  ,"Predictors of depression across samples")

```{r forest plot by sample no explan pcl, fig.height=30, fig.width=50, dpi=120}

lm.forest(plot.dat.noexplan.sample.pcl  ,"Predictors of symptoms of PTSD across samples")

### 2021-01-11 PANCHANGE_data_preprocessing2020-07-15.Rmd
Added keyworker status to employment variable
knitted exclusion file with all datasets updated for employment
detail checks and fixes

# Specify cohort
```{r Specify cohort}
GLAD = FALSE
EDGI = FALSE
NBR = FALSE
RAMP = TRUE
GLAD_and_EDGI = FALSE
```

@@

>>>>

# Specify cohort
```{r Specify cohort}
GLAD = FALSE
EDGI = FALSE
NBR = FALSE
RAMP = FALSE
GLAD_and_EDGI = TRUE
```

@@

# IMPUTE RETROSPECTIE DATA for no change

```{r impute retro}
# impute retro for those who did not change
dat$gad.sum_score_retro <- ifelse(dat$gad.pandemic_felt_feelings == "No different", dat$gad.sum_score_base,dat$gad.sum_score_retro)
dat$phq.sum_score_retro <- ifelse(dat$phq.pandemic_felt_feelings == "No different", dat$phq.sum_score_base,dat$phq.sum_score_retro)
```
